---
title: "Linear Regression"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = T, include = T, wwarning = F)
```

## Boston data

Boston is a database with information of areas around Boston city, and the median house prices. It is included in the MASS R package.
The essential data manipulation will be performed with the tidyverse collection of R packages.

```{r}
library(MASS)
library(tidyverse)
library(gridExtra)
library(broom)
```

We will use linear regression to infer the relationship between the percentage of lower status of the population (lstat) and house prices (medv).

## First linear regression model

Fit of the basic linear regression model using lm function. 

```{r}
fit1 <- lm(medv ~ lstat, data = Boston)
summary(fit1)
```
## Linear algebra

First look at the model matrix
```{r}
# Residuals from regression
e <- as.matrix(residuals(fit1))
e[1:10]
# Coefficients of the model matrix
X <- model.matrix(fit1) # from lm object
X[1:10,]

X <- cbind(1, Boston$lstat) # from dataset
X[1:10,]

Y <- Boston$medv
Y[1:10]
```


We now replicate exactly the coefficients estimated with the lm function by solving ordinary least square equations:

$$ \beta = (X^TX)^{-1}X^TY$$

```{r}
beta <- solve(t(X) %*% X) %*% t(X) %*% Y
beta
```

We now compute the H matrix.

$$ H = X(X^TX)^{-1}X^T$$

We check how the multiplication of $Y$ by $H$ give us the fitted values $\widehat Y$.

```{r}
H <- X %*% solve(t(X) %*% X) %*% t(X)
dim(H)
cbind(fitted(fit1), H %*% Y)[1:10,]
cbind(hatvalues(fit1), diag(H))[1:10,] 
cbind(
  e,                                         
  Y - X %*% beta,                            
  Y - H %*% Y,
  (diag(length(Y)) - H) %*% Y                   
)[1:10,]
```